### 2010-11-18 article.tex
git init

\documentclass[preprint,authoryear,12pt]{elsarticle}

\usepackage{amssymb}

\journal{Computers \& Chemical Engineering}

\begin{document}

%% ==== HEAD ====
\begin{frontmatter}

\title{A Systematic Approach to Model Predictive Controller Constraint Handling: Rigorous Geometric Methods}

\author{Andr\'e H. Campher\corref{cor1}}
\ead{ahcampher@tuks.co.za}

\author{Carl Sandrock}
%\ead{carl.sandrock@up.ac.za}

\cortext[cor1]{Corresponding author}

\address{Department of Chemical Engineering, University of Pretoria, Pretoria 0001, South Africa}

\begin{abstract}
some abstract
\end{abstract}

\begin{keyword}
model predictive control \sep whatever
\end{keyword}

\end{frontmatter}

% \linenumbers

%% ==== BODY ====
\section{Introduction}\label{sec:intro}
some intro
\citep{qhull}
%% ==== TAIL ====
\bibliographystyle{model4-names}
\bibliography{articlerefs}

\end{document}